---
title: "Hunters Green House Price prediction"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r houseprices,echo=FALSE}

library(tidyverse)
library(readxl)
library(dplyr)
library(glue)
```

```{r,echo=FALSE}
#options(scipen=0)
```

## Importing the data

```{r reading, echo=FALSE}
df <- read_xlsx("C:/College/Courses/SDM/Assignments/HuntersGreenHomeSales.xlsx", sheet = "Data")
```

## Data cleaning & Transformations

Made changes to 4 garages column values by getting values from Zillow

Making a copy of the original dataset
```{r}
df_new <- df
```

Renamed adom and cdom columns
```{r}
df_new <- rename(df_new, adom = adom_agentdaysonmarket)
df_new <- rename(df_new, cdom = cdom_cumuldaysmls)
```

Some of houses have been relisted and sold again between 2015 and 2020.

```{r}
df_new %>% 
  count(Address) %>% 
  filter(., n>1)

```

Taking only one record of those houses by filtering for latest selling date.
```{r}
df_new <- df_new %>% 
  mutate(Date = as.Date(x = datesold, format = "%y/%m/%d")) %>% 
  group_by(Address) %>% 
  arrange(desc(Date)) %>%
  slice(1) 
  
```

Checking for null values

```{r}
colSums((is.na(df_new)))
```

Replacing null values in Spa field with FALSE

```{r}
df_new <- df_new %>%
mutate(spa = replace_na(spa,FALSE))
```

### Data transformation

Halving the values of half baths and adding it to bathsfull

```{r}
df_new$bath <- (df_new$bathsfull + (df_new$bathshalf/2))
```

Finding how old a house was at the time the house was by finding the difference between year built and pending date.

```{r}
df_new$yrsold <- as.numeric(format(df_new$PendingDate, format = "%Y")) - df_new$yrblt
```

Changing values in roof field

```{r}
df_new <- df_new %>% 
  mutate(Roof = case_when(
    Roof == "Slate, Tile" ~ "Tile",
    Roof == "Shingle, Tile" ~ "Tile",
    Roof == "Shake, Shingle" ~ "Shingle",
    Roof == "Slate" ~ "Tile",
    Roof == "Built-Up" ~ "Shingle",
    Roof == "Other" ~ "Tile",
    Roof == "Concrete, Tile" ~ "Tile",
    Roof == "Tile" ~ "Tile",
    Roof == "Shingle" ~ "Shingle"
  ))
 
```

Count & percentage of categories in Roof

```{r}
janitor::tabyl(df_new, Roof)
janitor::tabyl(df_new, spa)
janitor::tabyl(df_new, Pool)
janitor::tabyl(df_new, splsale)
```

factorizing all categorical variables

```{r}
df_new$Roof <- factor(df_new$Roof)
df_new$Pool <- factor(df_new$Pool)
df_new$spa <- factor(df_new$spa)
df_new$splsale <- factor(df_new$splsale)
```

transferring cleaned and transformed dataframe to original one

```{r}
df <- df_new
remove(df_new)
```

removing columns which are not useful for predicting house prices

```{r}
df <- subset(x = df,select = -c(bathsfull,bathshalf,bathstotal,Address,Status,slnoskm,Date,datesold,sppersqft,subdivn))
```


### EDA

Histogram of dependent variables price sold and adom

```{r,echo=FALSE}
#require(gridExtra)
plot1 <- ggplot(data = df,aes(x=pricesold)) + geom_histogram()

plot2 <- ggplot(data = df,aes(x=log(pricesold))) + geom_histogram(bins=30)

gridExtra::grid.arrange(plot1, plot2, ncol=2)


```
The histogram is clearly skewed to the right. However doing a log on price sold makes it look normally distributed. We may have to use a exponential model for the price sold. 


Using the library performance analytics on numeric columns. This is to check for the relationship between all numeric variables.

```{r,echo=FALSE}
nums <- unlist(lapply(df, is.numeric))
PerformanceAnalytics::chart.Correlation(df[,nums])
```

We will not use variables in our model where correlation is greater than 0.7


```{r}
summary(df)
```

Does having a swimming pool affect price sold?

```{r,echo=FALSE}
df %>% 
  add_count(Pool,name={"Pool_total"}) %>% 
  mutate(Pool = glue("{Pool}({Pool_total})"),       fct_reorder(Pool,pricesold)) %>% 
  ggplot(aes(x = Pool,y=pricesold)) +
    geom_boxplot() 
           
```

Does type of Roof matter?

```{r,echo=FALSE}
ggplot(data = df, mapping = aes(x = Roof,y=pricesold)) +
    geom_boxplot()

```


```{r,echo=FALSE}
df %>% 
  ggplot(aes(pricesold,fill = Roof)) +
    geom_density(alpha = 0.5) 
```

Do old houses go for cheap price?

```{r,echo=FALSE}
df %>% 
  add_count(yrblt,name={"Year_total"}) %>% 
  mutate(yrblt = glue("{yrblt}({Year_total})"),       fct_reorder(yrblt,pricesold)) %>% 
  ggplot(aes(x = yrblt,y=pricesold)) +
    geom_jitter() + geom_boxplot(alpha=0.5)

#add alpha for transparency
```
```{r,echo=FALSE}
options(scipen = 999)
```

Spa's increase the value of a house?

```{r,echo=FALSE}
df %>% 
  ggplot(data = df, mapping = aes(x = spa,y=pricesold)) +
    geom_boxplot()
```

Does foreclosures affect selling price?

```{r,echo=FALSE}
df %>% 
  ggplot(data = df, mapping = aes(x = splsale,y=pricesold)) +
    geom_boxplot()
  
```

Having more beds increases house value?

```{r,echo=FALSE}
df %>% 
  ggplot(data = df, mapping = aes(x = Beds,y=pricesold)) +
    geom_point() + geom_smooth(method= "lm")
```


### Models

Exponential models predicting house value

```{r}
m1_sp <- lm(log(pricesold) ~ Beds + bath + garages + yrsold +  adom + Pool + spa + lotsqft + sqft + Roof, data = df)
summary(m1_sp)
plot(m1_sp)
```

Testing for OLS assumptions
```{r}
hist(m1_sp$residuals)
hist(m1_sp$fitted.values)

print("Normality - Shapiro Wilks test : ") # since sample less than 2000
shapiro.test(m1_sp$residuals) 
print("Bartlett's test : ") 
bartlett.test(list(m1_sp$residuals,m1_sp$fitted.values))
print("Durbin-Watson's test : ") 
lmtest::dwtest(m1_sp)
```

```{r}
#install.packages("car")
car::vif(m1_sp)
```

```{r}
m2_sp <- lm(log(pricesold) ~ Beds + bath + garages + yrsold + adom + Pool + spa + lppersqft + Roof, data = df)
summary(m2_sp)
plot(m2_sp)

hist(m2_sp$residuals)
hist(m2_sp$fitted.values)

print("Normality - Shapiro Wilks test : ") # since sample less than 2000
shapiro.test(m2_sp$residuals) 
print("Bartlett's test : ") 
bartlett.test(list(m2_sp$residuals,m2_sp$fitted.values))
print("Durbin-Watson's test : ") 
lmtest::dwtest(m2_sp)
car::vif(m2_sp)
#create years old field
```



```{r}
m3_sp <- lm(log(pricesold) ~ Beds + bath + garages + Pool + spa + lppersqft + Roof, data = df)
summary(m3_sp)
plot(m3_sp)

hist(m3_sp$residuals)
hist(m3_sp$fitted.values)

print("Normality - Shapiro Wilks test : ") # since sample less than 2000
shapiro.test(m3_sp$residuals) 
print("Bartlett's test : ") 
bartlett.test(list(m3_sp$residuals,m3_sp$fitted.values))
print("Durbin-Watson's test : ") 
lmtest::dwtest(m3_sp)
car::vif(m3_sp)
```

Comparing the three models to check our estimates are consistent.

```{r}
stargazer::stargazer(m1_sp,m2_sp,m3_sp, title="Results", type = "text")
```

Top 3 predictors of price sold from m2_sp:

* The marginal effect of baths suggests that adding 1 bathroom increases the selling price of the house by 20.44% keeping all other variables constant.
Private pool	
* The marginal effect of categorical variable Pool suggests that selling price for houses with private pool will be 9.85% more than houses with access to community pool.
* The marginal effect of the variable bed suggests that adding 1 bedroom increases the selling price of a house by 9.19% keeping all other variables constant.
